---
title: "p8105_hw3_xt2288"
output: github_document
date: "`r Sys.Date()`"
---

```{r}
library(tidyverse)
```


## Problem 1

```{r}
library(p8105.datasets)
data("instacart")
```

The data `instacart` has `r nrow(instacart)` observations and `r ncol(instacart)` variables. Some key variables are `order_id`: order identifier, `user_id`: customer identifier, `order_number`: the order sequence number for this user (1=first, n=nth), `order_dow`: the day of the week on which the order was placed, `order_hour_of_day`: the hour of the day on which the order was placed, `product_name`: name of the product, `aisle_id`: aisle identifier, `aisle`: the name of the aisle. Order_id 1 is made by user_id 112108. The 4th order is made on the 4th day of the week at 10. The product Bulgarian Yogurt is on aisle_id 120 which is yogurt.

* How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart |>
  group_by(aisle_id) |>
  summarize(n_obs = n()) |>
  nrow()

instacart |>
  group_by(aisle_id) |>
  summarize(n_obs = n()) |>
  filter(min_rank(desc(n_obs)) == 1)
```

There are 134 aisles, the most items are ordered from the 83 aisle.


* Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart |>
  group_by(aisle) |>
  summarize(n_order = n()) |>
  filter(n_order > 10000) |>
  arrange(n_order) |>
  mutate(aisle=factor(aisle, levels=aisle)) |>
  ggplot(aes(x = n_order, y = aisle)) +
  labs(
    title = "Number of items ordered in each aisle",
    xlab = "Number of items ordered",
    ylab = "Aisle"
  ) +
    geom_col()
```


* Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart |>
  filter(aisle == "baking ingredients" |aisle ==  "dog food care" | aisle == "packaged vegetables fruits") |>
  group_by(aisle, product_name) |>
  summarize(n_order = n()) |>
  mutate(pop_ranking = min_rank(desc(n_order))) |>
  filter(pop_ranking < 2) |>
  select(-pop_ranking)
```


* Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart |>
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") |>
  select(product_name,  order_dow, order_hour_of_day) |>
  group_by(product_name, order_dow) |>
  summarize(
    mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = "product_name", 
    values_from = "mean_hour")
```


## Problem 2

```{r}
data("brfss_smart2010")
```

data cleaning
```{r}
brfss_clean = brfss_smart2010 |>
  janitor::clean_names() |>
  filter(topic == "Overall Health") |>
  mutate(
    response = factor(response, levels=c("Poor","Fair","Good", "Very good","Excellent"))
  )
```

* In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r}
brfss_2002 = 
  brfss_clean |>
  filter(year == 2002) |>
  group_by(locationabbr, response) |>
  summarize(
    n_obs = n()
  ) |>
  filter(n_obs >=7)


brfss_2010 = 
  brfss_clean |>
  filter(year == 2010) |>
  count(locationabbr, name = "n_obs") |>
  filter(n_obs >= 7*5)
```

In 2002, `r unique(brfss_2002$locationabbr)` were observed at 7 or more locations. In 2010, `r brfss_2010$locationabbr` were observed at 7 or more locations.


* Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
brfss_clean |>
  filter(response == "Excellent") |>
  group_by(locationabbr, year) |>
  summarize(
    ave_data_value = mean(data_value) 
  ) |>
  ggplot(aes(x = year, y = ave_data_value, color = locationabbr)) +
    geom_line()
```


* Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
brfss_clean |>
  filter(locationabbr == "NY" & (year == 2006 | year == 2010)) |>
  ggplot(aes(x = response, y = data_value, color = year)) + 
    geom_boxplot() +
    facet_grid(. ~ year)
```


## Problem 3

```{r}
nhanes_covar = 
  read_csv("data/nhanes_covar.csv", skip = 4) |>
  filter(age >= 21) |>
  drop_na() |>
  mutate(
    sex = case_match(
      sex,
      1 ~ "male",
      2 ~ "female"),
    education = case_match(
      education,
      1 ~ "Less than high school",
      2 ~ "High school equivalent",
      3 ~ "More than high school"),
    education = factor(education, levels = c("Less than high school", "High school equivalent", "More than high school"))
  )
```

```{r}
nhanes_accel = 
  read_csv("data/nhanes_accel.csv") |>
  pivot_longer(
    min1:min1440,
    names_to = "minute",
    values_to = "MIMS_Value"
  )
```

```{r}
nhanes_merge = left_join(nhanes_covar, nhanes_accel, by = "SEQN")
```


a table for the number of men and women in each education category.
```{r}
nhanes_covar |>
  group_by(education, sex) |>
  summarize(
    n_obs = n()
  ) |>
  pivot_wider(
    names_from = sex,
    values_from = n_obs
  )
```
There is no big different for the number of male and female in less than high school and more than high school levels. 1 more female in less than high school level, and 3 more female in more than high school level. In the high school equivalent level, there are 12 more male than female.

a visualization of the age distributions for men and women in each education category
```{r}
nhanes_merge |>
  ggplot(aes(x = education, y = age, color = sex)) + 
  geom_boxplot() +
  labs(
    title = "Age vs. Education by Sex",
    x = "Education",
    y = "Age"
  )
```
In less than high school level, the age distribution is about the same for both sex. In high school equivalent level, female has a greater age than male. In more than high school level, male has a slightly greater age.


Plot total activities (y-axis) against age (x-axis); the plot compare men to women and have separate panels for each education level.
```{r}
nhanes_merge |>
  group_by(SEQN) |>
  mutate(total_act = sum(MIMS_Value)) |>
  filter(minute == "min1") |>
  ggplot(aes(x = age, y = total_act, color = sex)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ education) +
  labs(
    title = "Total Activity vs. Age",
    x = "Age",
    y = "Total Activiy"
  )
```
The total activity decreases as the age increases for both men and women in all education levels. In less than high school education level, women has a higher total activity before 40, and men has a higher activity after 40. In the high school equivalent level, women has a higher total activity. In more than high school level, female also has ahigher total activity. 


a three-panel plot that shows the 24-hour activity time courses for each education level and sex
```{r}
nhanes_merge |>
  mutate(minute = str_remove_all(minute, "min"),
         minute = as.numeric(minute)) |>
  arrange(minute) |>
  ggplot(aes(x = minute, y = MIMS_Value, color = sex)) +
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE) +
  facet_grid( ~ education) +
  labs(
    title = "MIMS_Value vs. Minute by Sex",
    x = "Minute",
    y = "MIMS Value"
  )
```
Based on the plots, women and men have about the same trend of 24-hour activity in all three education levels.








