---
title: "p8105_hw3_xt2288"
output: github_document
date: "`r Sys.Date()`"
---

```{r}
library(tidyverse)
```


## Problem 1

```{r}
library(p8105.datasets)
data("instacart")
```

The data `instacart` has `r nrow(instacart)` observations and `r ncol(instacart)` variables. Some key variables are `order_id`: order identifier, `user_id`: customer identifier, `order_number`: the order sequence number for this user (1=first, n=nth), `order_dow`: the day of the week on which the order was placed, `order_hour_of_day`: the hour of the day on which the order was placed, `product_name`: name of the product, `aisle_id`: aisle identifier, `aisle`: the name of the aisle. Order_id 1 is made by user_id 112108. The 4th order is made on the 4th day of the week at 10. The product Bulgarian Yogurt is on aisle_id 120 which is yogurt.

* How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart |>
  group_by(aisle_id) |>
  summarize(n_obs = n()) |>
  nrow()

instacart |>
  group_by(aisle_id) |>
  summarize(n_obs = n()) |>
  filter(min_rank(desc(n_obs)) == 1)
```

There are 134 aisles, the most items are ordered from the 83 aisle.


* Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart |>
  group_by(aisle) |>
  summarize(n_order = n()) |>
  filter(n_order > 10000) |>
  arrange(n_order) |>
  mutate(aisle=factor(aisle, levels=aisle)) |>
  ggplot(aes(x = n_order, y = aisle)) +
  labs(
    title = "Number of items ordered in each aisle",
    xlab = "Number of items ordered",
    ylab = "Aisle"
  ) +
    geom_col()
```


* Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart |>
  filter(aisle == "baking ingredients" |aisle ==  "dog food care" | aisle == "packaged vegetables fruits") |>
  group_by(aisle, product_name) |>
  summarize(n_order = n()) |>
  mutate(pop_ranking = min_rank(desc(n_order))) |>
  filter(pop_ranking < 2) |>
  select(-pop_ranking)
```


* Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart |>
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") |>
  select(product_name,  order_dow, order_hour_of_day) |>
  group_by(product_name, order_dow) |>
  summarize(
    mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = "product_name", 
    values_from = "mean_hour")
```


## Problem 2

```{r}
data("brfss_smart2010")
```

data cleaning
*format the data to use appropriate variable names?*
```{r}
brfss_clean = brfss_smart2010 |>
  janitor::clean_names() |>
  filter(topic == "Overall Health") |>
  mutate(
    response = factor(response, levels=c("Poor","Fair","Good", "Very good","Excellent"))
  )
```

* In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r}
brfss_2002 = 
  brfss_clean |>
  filter(year == 2002) |>
  group_by(locationabbr, response) |>
  summarize(
    n_obs = n()
  ) |>
  filter(n_obs >=7)


brfss_2010 = 
  brfss_clean |>
  filter(year == 2010) |>
  count(locationabbr, name = "n_obs") |>
  filter(n_obs >= 7*5)
```

In 2002, `r unique(brfss_2002$locationabbr)` were observed at 7 or more locations. In 2010, `r brfss_2010$locationabbr` were observed at 7 or more locations.


* Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
brfss_clean |>
  filter(response == "Excellent") |>
  group_by(locationabbr, year) |>
  summarize(
    ave_data_value = mean(data_value) 
  ) |>
  ggplot(aes(x = year, y = ave_data_value, color = locationabbr)) +
  geom_line()
```


* Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.







